---
title: "C.P.E. Bach"
author: "Herma Ardesch"
date: "13-2-2021"
output: html_document
---

## Authentic vs. Modern Performances

This portfolio is designed to show statistical data on the research question: *Is it possible to detect differences between authentic and modern performances in the music of C.Ph.E. Bach, using the Spotify API?* To give a first impression, the next graphs are added.

### Tempo

```{r}

library(tidyverse)
library(spotifyr)

CPE_aut <- get_playlist_audio_features("", "2PmIxISPPF4ymvoa3bvCrh")
CPE_mod <- get_playlist_audio_features("", "0F83MSl8bV2AxL2JNDTGew")
CPE <-
  bind_rows(
    CPE_aut %>% mutate(category = "authentic"),
    CPE_mod %>% mutate(category = "modern")
  )
CPE %>% ggplot(aes(x = tempo)) + 
 geom_histogram(binwidth = 10, color = "grey80", fill = "grey60") +
  facet_wrap(~category) +
labs(title = "Differences in tempo of authentic vs modern performances", caption = "Source: Spotify API")


```



The two histograms show the tempo variances between authentic and modern performances of music by Carl Philipp Emanuel Bach, each playlist consisting of 45 identical movements. The histogram on the left shows a more even pattern with less variances, whereas the modern performances are more outspoken in tempo. TO DO: explain peak at 80!

### Add title

```{r}

CPE %>% ggplot(aes(x = tempo, y = acousticness, color = category)) +
  geom_jitter() + geom_smooth() +
labs(title = "Differences in timbre of authentic vs modern performances", caption = "Source: Spotify API")

```


This scatter plot shows the acousticness vs tempo of both groups, authentic and modern. TO DO: change variables!
