---
title: "C.P.E. Bach"
author: "Herma Ardesch"
date: "13-2-2021"
output: 
  flexdashboard::flex_dashboard:
   storyboard: true
   theme: yeti
   vertical layout: fill
---

## Authentic vs. Modern Performances of C.P.E. Bach's music

### **Why C.P.E. Bach?** Discover more about the background of this portfolio...

```{r}

```

> 

#### **The music of Carl Philipp Emanuel Bach** ####

Raised in an environment in which Bach's music was predominant, I learned to play the flute and developed a preference for the music of Carl Philipp Emanuel Bach (1714_1788). His 'Empfindsame Stil' with many unexpected melodic, harmonic and thythmic turns to symbolize mood swings was very appealing to me. This corpus consists of his flute concertos, cello concertos, some of his harpsichord sonatas and his four symphonies.

I would like to answer the question **Is it possible to detect differences between authentic and modern performance in the music of C.Ph.E. Bach, using the Spotify API?** I expect differences in tempo, interpretation, instrumental sound, pitch, valence and energy between authentic and modern performance practices. Of course I need to give extra attention to the pitch of both groups, since authentic instruments usually play in a different tuning, resulting in a lower pitch. My objective it to make these differences visible as well as audible.

Listening to his Flute Concerto in A minor, Wq 166 in the authentic performance of Konrad Hünteler with the Amsterdam Baroque Orchestra (Ton Koopman) and the modern performance of Emmanuel Pahud with the Kammerakademie Potsdam (Trevor Pinnock), you can hear the diffenrences. Both performances take a different tempo, energy level and pitch and the wooden baroque traverso has quite a different sound character, compared to the modern silver or golden Böhm-flute. 

However, I expect that since the inception of the authentic performance practice over half a century ago, both authentic and modern performances have grown closer to each other, because of improved replicas of authentic instruments on the one hand and a greater versatility of instrumentalists playing both authentic and modern versions of their instrument.
 

### **See what you hear?** Have a look at a visualisation of **TEMPO** differences**...

```{r}

library(tidyverse)
library(spotifyr)
library(plotly)

CPE_aut <- get_playlist_audio_features("", "2PmIxISPPF4ymvoa3bvCrh")
CPE_mod <- get_playlist_audio_features("", "0F83MSl8bV2AxL2JNDTGew")
CPE <-
  bind_rows(
    CPE_aut %>% mutate(category = "Authentic"),
    CPE_mod %>% mutate(category = "Modern")
  )
tempo_gg <- CPE %>% ggplot(aes(x = tempo)) + 
 geom_histogram(binwidth = 10, color = "grey80", fill = "deepskyblue4") +
  facet_wrap(~category) +
labs(title = "Differences in tempo of authentic vs modern performances", size = 16)

ggplotly(tempo_gg)

```

> Source: Spotify API

*** 

Moving your cursor over the top of the histograms will show you an interactive menu. You can zoom in on the bins to see how often each tempo bin is used.

### **See what you hear?** A first impression of **TIMBRE differences**...

```{r}

CPE %>% ggplot(aes(x = tempo, y = acousticness, color = category)) +
  geom_jitter() + geom_smooth() +
labs(title = "Differences in timbre of authentic vs modern performances")
```

> Source: Spotify API

***

This scatter plot shows the acousticness vs tempo of both groups, authentic and modern. TO DO: change variables!
